KidsAIR: merge exposure datasets

### 2020-06-26 merge_datasets.Rmd
data cleaning/merging updates

---
title: "KidsAIR: merge exposure datasets"
author: "Ethan Walker"
date: "Started 24 March 2020, Updated 14 April 2020"
output: pdf_document
---

@@

# Load individual datasets
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

pm_clean <- 
  read_rds(paste0(output_path, "Output/pm_clean.rds")) 
sums_clean <- 
  read_rds(paste0(output_path, "Output/sums_clean.rds"))

kids_linked_ids <- 
  read_rds(paste0(output_path, "Output/kids_linked_ids.rds"))
demographics_clean <- 
  read_rds(paste0(output_path, "Output/demographics_clean.rds")) 
moisture_clean <- 
  read_rds(paste0(output_path, "Output/moisture_clean.rds"))
indoor_temp_rh_clean <- 
  read_rds(paste0(output_path, "Output/indoor_temp_rh_clean.rds"))
stove_use_clean <- 
  read_rds(paste0(output_path, "Output/stove_use_clean.rds"))
stove_grades_clean <- 
  read_rds(paste0(output_path, "Output/stove_grades_clean.rds")) 
home_char_clean <- 
  read_rds(paste0(output_path, "Output/home_char_clean.rds"))
home_act_clean <- 
  read_rds(paste0(output_path, "Output/home_act_clean.rds"))
```

# Create new dataset with mean PM/SUMs; join with covariates
## Long data averaged over sampling date, including SUMs dates, repeated by child_id
```{r}
# Prep PM data to join with SUMs data by sampling day
pm_join <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ",
           remove = FALSE) %>% 
  group_by(area, home, sample_date) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE),
         pm_daily_obs = n()) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, dtid, day_of_week, 
         pm_total_observations, pm_mean_sampling_period, sampling_day, 
         pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
         sample_date, pm_mean_daily, pm_daily_obs, pm_spike_1000,
         pm_spike_2000, pm_spike_5000, pm_sample_interval) %>% 
  group_by(area, home) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

#write_rds(pm_join, paste0(output_path, "Output/pm_join.rds"))
pm_join <- read_rds(paste0(output_path, "Output/pm_join.rds"))

# Prep SUMs data to join with PM data
sums_join <- sums_clean %>% 
  mutate(sample_date = as.character(date_sums)) %>% 
  select(area, home, home_winter_id, winter_id, ibid, treatment, 
         datetime_sums, sample_date, temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(sums_mean_winter = mean(temp_c_sums, na.rm = TRUE),
         sums_mean_5 = sums_mean_winter/5) %>% 
  group_by(area, home, winter_id, sample_date) %>% 
  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
         temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0),
         temp_perc_30 = (sum(temp_perc_30)/n()*100),
         temp_perc_27 = if_else(temp_c_sums >= 27, 1, 0),
         temp_perc_27 = (sum(temp_perc_27)/n()*100),
         temp_perc_25 = if_else(temp_c_sums >= 25, 1, 0),
         temp_perc_25 = (sum(temp_perc_25)/n()*100),
         sums_daily_obs = n()) %>% 
  #distinct(sums_mean_daily, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

#write_rds(sums_join, paste0(output_path, "Output/sums_join.rds"))
sums_join <- read_rds(paste0(output_path, "Output/sums_join.rds"))

# Join PM/SUMs data and filter/arrange by sampling day
pm_sums_joined1 <- sums_join %>% 
  left_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  arrange(area, home, datetime_sums) %>% 
  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
  filter(filter_var == 1) %>% 
  group_by(area, home, winter_id) %>%
  mutate(sums_mean_sampling_period = mean(temp_c_sums, na.rm = TRUE)) %>% 
  group_by(area, home, winter_id, sampling_day) %>% 
  mutate(sums_mean_sampling_day = mean(temp_c_sums, na.rm = TRUE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day, 
         sums_mean_sampling_day, sums_mean_sampling_period)

#write_rds(pm_sums_joined1, paste0(output_path, "Output/pm_sums_joined1.rds"))
pm_sums_joined1 <- read_rds(paste0(output_path, "Output/pm_sums_joined1.rds"))

pm_sums_joined2 <- pm_join %>% 
  left_join(pm_sums_joined1, 
            by = c("area", "home", "home_winter_id", "winter_id", "treatment", "sampling_day"))

# Join SUMs data with original PM data, by sampling day
pm_sums_joined3 <- sums_join %>% 
  select(-datetime_sums, -temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  arrange(area, home, winter_id, sample_date) %>% 
  full_join(pm_sums_joined2, by = c("area", "home", "home_winter_id", "winter_id", 
                                    "treatment", "sample_date")) %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

# Join PM/SUMs data with other covariate datasets
analysis_data1 <- home_act_clean %>% 
  mutate(sampling_day = home_act_day) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  select(-home_act_day, -home_act_comments, -home_id) %>% 
  full_join(pm_sums_joined3, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment", "sampling_day")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id, sample_date)

analysis_data2 <- demographics_clean %>% 
  mutate(residents_under_five = if_else(residents_under_five == 0 & age_child_nov != 4, "NA", residents_under_five),
         residents_under_five = if_else(residents_under_five == 0 & child_id_char == "WMA673B", "NA", residents_under_five)) %>% 
  replace_with_na(replace = list(total_residents = "0")) %>% 
  replace_with_na(replace = list(total_residents = "1")) %>% 
  replace_with_na(replace = list(residents_under_five = "NA")) %>% 
  full_join(analysis_data1, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  mutate(income_3level = fct_collapse(income,
                               "Less than $20,000" = "Less than $20,000",
                               "$20,000-$49,999" = 
                                 c("$20,000 to $29,999", "$30,000 to $39,999",
                                   "$40,000 to $49,999"),
                                "$50,000+" = 
                                 c("$50,000 to $74,999", "$75,000 to $99,999",
                                   "$100,000 or more")),
         income_3level = factor(income_3level, levels = c("Less than $20,000", 
                                            "$20,000-$49,999", 
                                            "$50,000+")), 
         education_3level = fct_collapse(education,
                                  "High school or less" = 
                                    c("Less than high school", 
                                      "High school diploma or GED"),
                                  "Some college" = "Some college",
                                  "College degree" = "College degree"),
         education_3level = factor(education_3level, levels = c("High school or less", 
                                            "Some college", 
                                            "College degree")), 
         race_parent_2level = fct_collapse(race_parent,
                                    "White" = "White",
                                    "Non-white" = 
                                      c("American Indian/Alaskan Native",
                                        "Asian", "More than one race")),
         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
         total_residents_2level = fct_collapse(total_residents,
                                    "<5" = c("1", "2", "3", "4"),
                                    "5+" = c("5", "6", "7", "8", "9", "10",
                                             "11", "12", "13", "14", "15")),
         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
         residents_under5_2level = fct_collapse(residents_under_five,
                                    "1" = c("0", "1"),
                                    "2+" = c("2", "3", "4", "5")),
         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

@@

  select(area, home, home_winter_id, winter_id, treatment, residents_smoke:home_insert) %>% 
  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 

@@

analysis_data4 <- indoor_temp_rh_clean %>% 
  filter(winter_id == 1) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment, rh_indoor_current:temp_indoor_min) %>% 
  ungroup() %>% 
  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", "treatment")) %>% 
  mutate(temp_max_5 = temp_indoor_max/5,
         temp_min_5 = temp_indoor_min/5,
         rh_max_5 = rh_indoor_max/5,
         rh_min_5 = rh_indoor_min/5) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

@@

pm_moisture <- moisture_clean %>% 
  mutate(moisture_date = as.character(moisture_date),
         winter_id = as.character(winter_id)) %>% 
  filter(!is.na(moisture_ave)) %>% 
  select(area, home, home_winter_id, winter_id, treatment, moisture_ave, 
         moisture_split, moisture_date) %>% 
  left_join(pm_join2, by = c("area", "home", "winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(mean_moisture_winter = mean(moisture_ave, na.rm = TRUE),
         moisture_obs_winter = n(),
         moisture_date = ymd(moisture_date),
         pm_first_date = ymd(pm_first_date),
         time_diff = moisture_date - pm_first_date,
         time_diff_rank = rank(abs(time_diff)),
         moisture_split_numeric = factor(moisture_split,
                                 levels = c("Yes", "No"),
                                 labels = c(1, 0)),
         moisture_split_numeric = as.numeric(moisture_split_numeric),
         moisture_split_numeric = if_else(moisture_split_numeric == 2, 0, moisture_split_numeric),
         moisture_split_percent_winter = (sum(moisture_split_numeric)/n())*100,
         moisture_split_closest = moisture_split) %>% 
  filter(time_diff_rank == 1) %>% 
  mutate(moisture_closest = if_else(!is.na(pm_first_date), moisture_ave, 999),
         moisture_closest_date = moisture_date,
         moisture_closest_date = as.character(moisture_closest_date),
         moisture_closest_date = if_else(is.na(time_diff), "NA", moisture_closest_date),
         pm_moisture_days_diff = time_diff) %>% 
  replace_with_na(replace = list(moisture_closest = 999)) %>% 
  replace_with_na(replace = list(moisture_closest_date = "NA")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id) %>% 
  select(-moisture_ave, -time_diff, -time_diff_rank, -moisture_split, 
         -moisture_split_numeric)

analysis_data5 <- pm_moisture %>% 
  right_join(analysis_data4, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data6 <- stove_grades_clean %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment, stove_grade) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  right_join(analysis_data5, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  mutate(stove_grade_3level = fct_collapse(stove_grade,
                                    "A" = "A",
                                    "C" = "C",
                                    "F" = c("D", "F"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data7 <- stove_use_clean %>% 
  filter(winter_id == 1) %>% 
  select(-home_id) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  right_join(analysis_data6, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment")) %>% 
  mutate(burn_level_3level = fct_collapse(burn_level,
                                   "No/Light burning" = c("No burning", 
                                                          "Light burning"),
                                   "Ave burning" = "Average burning",
                                   "Heavy burning" = "Heavy burning"),
         burn_level_3level = factor(burn_level_3level,
                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
         wood_collect_2level = fct_collapse(wood_collect,
                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
                                                   "1 mo to 3 mo"),
                                   "3+ months" = c("3 mo to 6 mo", 
                                                   "6 mo to 1 year", "> 1 yr")),
         wood_collect_2level = factor(wood_collect_2level,
                         levels = c("<3 months", "3+ months")),
         wood_collect_method_2level = fct_collapse(wood_collect_method,
                                   "Harvest yourself" = "Harvest yourself",
                                   "Purchase/other" = c("Mix", "Other", 
                                                        "Purchase it")),
         wood_cords = if_else(wood_cords > 10, 10, wood_cords)) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

write_rds(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.rds"))
write_csv(analysis_data7, paste0(output_path, "Output/exposure_analysis_data_long.csv"))

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))

@@

#write_csv(save_var_names, paste0(output_path, "Output/kids_data_var_names.csv"))
```

# Update long dataset -- REMEMBER TO UPDATE SHORT DATA BELOW
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long.rds"))

kids_amb_pm_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date)

kids_amb_temp_data <- 
  read_rds(paste0(output_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date))

@@

write_rds(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))
write_csv(exposure_analysis_data_long_new, paste0(output_path, "Output/exposure_analysis_data_long_new.csv"))
```

# Convert long data to medium: only keeping days with PM sampling, repeated by child_id
```{r}
output_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long_new <- 
  read_rds(paste0(output_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- exposure_analysis_data_long_new %>% 
  select(area, home, home_winter_id, winter_id, treatment, 
         sampling_day, sample_date, burn_level:mean_temp_roll_4day) %>% 
  group_by(area, home, winter_id, child_id_char) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  select(area, home, home_winter_id, winter_id, treatment, sampling_day,
         child_id_char, burn_level:mean_temp_roll_4day) %>% 
  mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
         filter_var = if_else(is.na(filter_var), 0, filter_var)) %>% 
  select(filter_var, area:mean_temp_roll_4day) %>% 
  filter(filter_var == 0) %>% 
  select(-filter_var) %>% 
  ungroup()

>>>>

---
title: "KidsAIR: merge exposure datasets"
author: "Ethan Walker"
date: "Started 24 March 2020, Updated 26 June 2020"
output: pdf_document
---

@@

# Load individual datasets
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

kids_linked_ids <- read_rds(paste0(file_path, "Output/kids_linked_ids_final.rds")) 
demographics <- read_rds(paste0(file_path, "Output/demographics_final.rds"))

pm_clean <- 
  read_rds(paste0(file_path, "Output/pm_clean.rds")) 
sums_clean <- 
  read_rds(paste0(file_path, "Output/sums_clean.rds"))

moisture_clean <- 
  read_rds(paste0(file_path, "Output/moisture_clean.rds"))
indoor_temp_rh_clean <- 
  read_rds(paste0(file_path, "Output/indoor_temp_rh_clean.rds"))
stove_use_clean <- 
  read_rds(paste0(file_path, "Output/stove_use_clean.rds"))
stove_grades_clean <- 
  read_rds(paste0(file_path, "Output/stove_grades_clean.rds")) 
home_char_clean <- 
  read_rds(paste0(file_path, "Output/home_char_clean.rds"))
home_act_clean <- 
  read_rds(paste0(file_path, "Output/home_act_clean.rds"))
```

# Create new dataset with mean PM/SUMs; join with covariates
## Long data averaged over sampling date, including SUMs dates, repeated by child_id
```{r}
# Prep PM data to join with SUMs data by sampling day
pm_join <- pm_clean %>% 
  group_by(area, home) %>% 
  mutate(pm_mean_sampling_period = mean(pm, na.rm = TRUE)) %>% 
  group_by(area, home, sampling_day) %>% 
  mutate(pm_mean_sampling_day = mean(pm, na.rm = TRUE)) %>% 
  mutate(sample_datetime = ymd_hms(pm_datetime_new)) %>% 
  separate(sample_datetime, c("sample_date", "sample_time"), sep = " ",
           remove = FALSE) %>% 
  group_by(area, home, sample_date) %>% 
  mutate(pm_mean_daily = mean(pm, na.rm = TRUE),
         pm_daily_obs = n()) %>% 
  group_by(area, home, sampling_day) %>% 
  arrange(sample_datetime) %>% 
  mutate(pm_last_datetime = last(sample_datetime),
         pm_first_datetime = first(sample_datetime)) %>% 
  mutate(winter_id = as.character(winter_id)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date, desc(pm_spike_1000),
          desc(pm_spike_2000), desc(pm_spike_5000)) %>% 
  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
         dtid, day_of_week, pm_total_observations, pm_mean_sampling_period, 
         sampling_day, pm_mean_sampling_day, pm_first_datetime, pm_last_datetime,
         sample_date, pm_mean_daily, pm_daily_obs, pm_spike_1000,
         pm_spike_2000, pm_spike_5000, pm_sample_interval) %>% 
  group_by(area, home) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

#write_rds(pm_join, paste0(file_path, "Output/pm_join.rds"))
pm_join <- read_rds(paste0(file_path, "Output/pm_join.rds"))

# Prep SUMs data to join with PM data
sums_join <- sums_clean %>% 
  mutate(sample_date = as.character(date_sums)) %>% 
  select(area, home, home_winter_id, winter_id, ibid, treatment_assigned, 
         treatment_actual, datetime_sums, sample_date, temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(sums_mean_winter = mean(temp_c_sums, na.rm = TRUE),
         sums_mean_5 = sums_mean_winter/5) %>% 
  group_by(area, home, winter_id, sample_date) %>% 
  mutate(sums_mean_daily = mean(temp_c_sums, na.rm = TRUE),
         temp_perc_30 = if_else(temp_c_sums >= 30, 1, 0),
         temp_perc_30 = (sum(temp_perc_30)/n()*100),
         temp_perc_27 = if_else(temp_c_sums >= 27, 1, 0),
         temp_perc_27 = (sum(temp_perc_27)/n()*100),
         temp_perc_25 = if_else(temp_c_sums >= 25, 1, 0),
         temp_perc_25 = (sum(temp_perc_25)/n()*100),
         sums_daily_obs = n()) %>% 
  #distinct(sums_mean_daily, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, datetime_sums)

#write_rds(sums_join, paste0(file_path, "Output/sums_join.rds"))
sums_join <- read_rds(paste0(file_path, "Output/sums_join.rds"))

# Join PM/SUMs data and filter/arrange by sampling day
pm_sums_joined1 <- sums_join %>% 
  left_join(pm_join, by = c("area", "home", "home_winter_id", "winter_id", 
                            "treatment_assigned", "treatment_actual")) %>% 
  arrange(area, home, datetime_sums) %>% 
  mutate(filter_var = if_else(datetime_sums >= pm_first_datetime &
                              datetime_sums <= pm_last_datetime, 1, 0)) %>% 
  filter(filter_var == 1) %>% 
  group_by(area, home, winter_id) %>%
  mutate(sums_mean_sampling_period = mean(temp_c_sums, na.rm = TRUE)) %>% 
  group_by(area, home, winter_id, sampling_day) %>% 
  mutate(sums_mean_sampling_day = mean(temp_c_sums, na.rm = TRUE)) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  ungroup() %>% 
  arrange(area, home, sampling_day) %>% 
  select(area, home, home_winter_id, winter_id, sampling_day, 
         sums_mean_sampling_day, sums_mean_sampling_period,
         treatment_assigned, treatment_actual)

#write_rds(pm_sums_joined1, paste0(file_path, "Output/pm_sums_joined1.rds"))
pm_sums_joined1 <- read_rds(paste0(file_path, "Output/pm_sums_joined1.rds"))

pm_sums_joined2 <- pm_join %>% 
  left_join(pm_sums_joined1, 
            by = c("area", "home", "home_winter_id", "winter_id", 
                   "treatment_assigned", "treatment_actual", "sampling_day"))

# Join SUMs data with original PM data, by sampling day
pm_sums_joined3 <- sums_join %>% 
  select(-datetime_sums, -temp_c_sums) %>% 
  group_by(area, home, winter_id) %>% 
  distinct(sample_date, .keep_all = TRUE) %>% 
  arrange(area, home, winter_id, sample_date) %>% 
  full_join(pm_sums_joined2, by = c("area", "home", "home_winter_id", "winter_id", 
                                    "treatment_assigned", "treatment_actual", 
                                    "sample_date")) %>% 
  mutate(sampling_day = as.character(sampling_day)) %>% 
  ungroup() %>% 
  arrange(area, home, sample_date)

# Join PM/SUMs data with other covariate datasets
analysis_data1 <- home_act_clean %>% 
  mutate(sampling_day = home_act_day) %>% 
  select(-home_act_day, -home_id) %>% 
  full_join(pm_sums_joined3, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment_assigned", 
                                    "treatment_actual", "sampling_day")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id, sample_date)

analysis_data2 <- demographics %>% 
  mutate(residents_under_five = if_else(residents_under_five == 0 & age_child_nov != 4, "NA", residents_under_five),
         residents_under_five = if_else(residents_under_five == 0 & child_id_char == "WMA673B", "NA", residents_under_five)) %>% 
  replace_with_na(replace = list(total_residents = "0")) %>% 
  replace_with_na(replace = list(total_residents = "1")) %>% 
  replace_with_na(replace = list(residents_under_five = "NA")) %>% 
  full_join(analysis_data1, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment_assigned", 
                                    "treatment_actual")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  mutate(income_3level = fct_collapse(income,
                               "Less than $20,000" = "Less than $20,000",
                               "$20,000-$49,999" = 
                                 c("$20,000 to $29,999", "$30,000 to $39,999",
                                   "$40,000 to $49,999"),
                                "$50,000+" = 
                                 c("$50,000 to $74,999", "$75,000 to $99,999",
                                   "$100,000 or more")),
         income_3level = factor(income_3level, levels = c("Less than $20,000", 
                                            "$20,000-$49,999", 
                                            "$50,000+")), 
         education_3level = fct_collapse(education,
                                  "High school or less" = 
                                    c("Less than high school", 
                                      "High school diploma or GED"),
                                  "Some college" = "Some college",
                                  "College degree" = "College degree"),
         education_3level = factor(education_3level, levels = c("High school or less", 
                                            "Some college", 
                                            "College degree")), 
         race_parent_2level = fct_collapse(race_parent,
                                    "White" = "White",
                                    "Non-white" = 
                                      c("American Indian/Alaskan Native",
                                        "Asian", "More than one race")),
         race_parent_2level = factor(race_parent_2level, levels = c("White", "Non-white")), 
         total_residents_2level = fct_collapse(total_residents,
                                    "<5" = c("1", "2", "3", "4"),
                                    "5+" = c("5", "6", "7", "8", "9", "10",
                                             "11", "12", "13", "14", "15")),
         total_residents_2level = factor(total_residents_2level, levels = c("<5", "5+")),
         residents_under5_2level = fct_collapse(residents_under_five,
                                    "1" = c("0", "1"),
                                    "2+" = c("2", "3", "4", "5")),
         residents_under5_2level = factor(residents_under5_2level, levels = c("1", "2+"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

@@

  select(area, home, home_winter_id, winter_id, treatment_assigned, 
         treatment_actual, residents_smoke:home_insert) %>% 
  right_join(analysis_data2, by = c("area", "home", "home_winter_id", "winter_id", 
                                    "treatment_assigned", "treatment_actual")) %>% 

@@

analysis_data4 <- indoor_temp_rh_clean %>% 
  filter(winter_id == 1) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
         rh_indoor_current:temp_indoor_min) %>% 
  ungroup() %>% 
  right_join(analysis_data3, by = c("area", "home", "home_winter_id", "winter_id", 
                                    "treatment_assigned", "treatment_actual")) %>% 
  mutate(temp_max_5 = temp_indoor_max/5,
         temp_min_5 = temp_indoor_min/5,
         rh_max_5 = rh_indoor_max/5,
         rh_min_5 = rh_indoor_min/5) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

@@

pm_moisture <- moisture_clean %>% 
  mutate(moisture_date = as.character(moisture_date),
         winter_id = as.character(winter_id)) %>% 
  filter(!is.na(moisture_ave)) %>% 
  select(area, home, home_winter_id, winter_id, treatment_actual, 
         treatment_assigned, moisture_ave, 
         moisture_split, moisture_date) %>% 
  left_join(pm_join2, by = c("area", "home", "winter_id")) %>% 
  arrange(area, home, winter_id) %>% 
  group_by(area, home, winter_id) %>% 
  mutate(mean_moisture_winter = mean(moisture_ave, na.rm = TRUE),
         moisture_obs_winter = n(),
         moisture_date = ymd(moisture_date),
         pm_first_date = ymd(pm_first_date),
         time_diff = moisture_date - pm_first_date,
         time_diff_rank = rank(abs(time_diff)),
         moisture_split_numeric = factor(moisture_split,
                                 levels = c("Yes", "No"),
                                 labels = c(1, 0)),
         moisture_split_numeric = as.numeric(moisture_split_numeric),
         moisture_split_numeric = if_else(moisture_split_numeric == 2, 0, moisture_split_numeric),
         moisture_split_percent_winter = (sum(moisture_split_numeric)/n())*100,
         moisture_split_closest = moisture_split) %>% 
  filter(time_diff_rank == 1) %>% 
  mutate(moisture_closest = if_else(!is.na(pm_first_date), moisture_ave, 999),
         moisture_closest_date = moisture_date,
         moisture_closest_date = as.character(moisture_closest_date),
         moisture_closest_date = if_else(is.na(time_diff), "NA", moisture_closest_date),
         pm_moisture_days_diff = time_diff) %>% 
  replace_with_na(replace = list(moisture_closest = 999)) %>% 
  replace_with_na(replace = list(moisture_closest_date = "NA")) %>% 
  ungroup() %>% 
  arrange(area, home, winter_id) %>% 
  select(-moisture_ave, -time_diff, -time_diff_rank, -moisture_split, 
         -moisture_split_numeric)

analysis_data5 <- pm_moisture %>% 
  right_join(analysis_data4, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment_assigned", 
                                    "treatment_actual")) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data6 <- stove_grades_clean %>% 
  arrange(area, home, winter_id) %>% 
  group_by(home) %>% 
  mutate(stove_grade = if_else(is.na(stove_grade), lead(stove_grade), stove_grade)) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  select(area, home, home_winter_id, winter_id, treatment_actual, 
         treatment_assigned, stove_grade) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  right_join(analysis_data5, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment_assigned", 
                                    "treatment_actual")) %>% 
  mutate(stove_grade_3level = fct_collapse(stove_grade,
                                    "A" = "A",
                                    "C" = "C",
                                    "F" = c("D", "F"))) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

analysis_data7 <- stove_use_clean %>% 
  filter(winter_id == 1) %>% 
  select(-home_id) %>% 
  distinct(home, .keep_all = TRUE) %>% 
  mutate(winter_id = as.character(winter_id)) %>%
  right_join(analysis_data6, by = c("area", "home", "home_winter_id", 
                                    "winter_id", "treatment_assigned", 
                                    "treatment_actual")) %>% 
  mutate(burn_level_3level = fct_collapse(burn_level,
                                   "No/Light burning" = c("No burning", 
                                                          "Light burning"),
                                   "Ave burning" = "Average burning",
                                   "Heavy burning" = "Heavy burning"),
         burn_level_3level = factor(burn_level_3level,
                         levels = c("No/Light burning", "Ave burning", "Heavy burning")),
         wood_collect_2level = fct_collapse(wood_collect,
                                   "<3 months" = c("< 1 week", "1 week to 1 mo",
                                                   "1 mo to 3 mo"),
                                   "3+ months" = c("3 mo to 6 mo", 
                                                   "6 mo to 1 year", "> 1 yr")),
         wood_collect_2level = factor(wood_collect_2level,
                         levels = c("<3 months", "3+ months")),
         wood_collect_method_2level = fct_collapse(wood_collect_method,
                                   "Harvest yourself" = "Harvest yourself",
                                   "Purchase/other" = c("Mix", "Other", 
                                                        "Purchase it")),
         wood_cords = if_else(wood_cords > 10, 10, wood_cords)) %>% 
  ungroup() %>% 
  arrange(area, home, home_winter_id, sample_date)

write_rds(analysis_data7, paste0(file_path, "Output/exposure_analysis_data_long.rds"))
write_csv(analysis_data7, 
          paste0(file_path, "Output/exposure_analysis_data_long.csv"), na = "")

exposure_analysis_data_long <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_long.rds"))

@@

#write_csv(save_var_names, paste0(file_path, "Output/kids_data_var_names.csv"))
```

# Update long dataset -- REMEMBER TO UPDATE SHORT DATA BELOW
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_long.rds"))

kids_amb_pm_data <- 
  read_rds(paste0(file_path, "Output/kids_amb_pm_data_new.rds")) %>% 
  mutate(sample_date = amb_pm_date)

kids_amb_temp_data <- 
  read_rds(paste0(file_path, "Output/kids_amb_temp_data_new.rds")) %>% 
  mutate(sample_date = temp_date) %>% 
  mutate(sample_date = as.character(sample_date))

@@

write_rds(exposure_analysis_data_long_new, paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))
write_csv(exposure_analysis_data_long_new, 
          paste0(file_path, "Output/exposure_analysis_data_long_new.csv"), na = "")
```

# Convert long data to medium: only keeping days with PM sampling, repeated by child_id
```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/KidsAIR/")

exposure_analysis_data_long_new <- 
  read_rds(paste0(file_path, "Output/exposure_analysis_data_long_new.rds"))

exposure_analysis_data_medium <- exposure_analysis_data_long_new %>% 
  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
         sampling_day, sample_date, burn_level:mean_temp_roll_4day) %>% 
  group_by(area, home, winter_id, child_id_char) %>% 
  distinct(sampling_day, .keep_all = TRUE) %>% 
  select(area, home, home_winter_id, winter_id, treatment_assigned, treatment_actual, 
         sampling_day, child_id_char, burn_level:mean_temp_roll_4day) %>% 
  mutate(filter_var = if_else(winter_id == 1 & is.na(sampling_day) & lead(sampling_day) == 1, 1, 0),
         filter_var = if_else(is.na(filter_var), 0, filter_var)) %>% 
  select(filter_var, area:mean_temp_roll_4day) %>% 
  filter(filter_var == 0) %>% 
  select(-filter_var) %>% 
  ungroup()
